---
title: "FiRAM"
subtitle: Fisheries Resource Allocation Model
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FiRAM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: Tidbury2021
  title: "Balancing biological and economic goals in commercial and recreational fisheries: systems modelling of sea bass fisheries"
  author: 
   - family: Tidbury
     given: Hannah J
   - family: Muench
     given: Angela
   - family: Lamb
     given: Philip D
   - family: Hyder
     given: Kieran

  container-title: ICES Journal of Marine Science
  volume: 78
  page: 1793-1803
  type: article-journal 
  issued: 
    year: 2021  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(FiRAM)
```
# Introduction 

* What is the model?
* What is this document?

# Setting up the model

* Loading in saved parameters
  + Changing some parameters
  + Maybe more??
* Defining custom `BioeconomicParams` objects from a file

In `FiRAM`, model parameters are contained in objects of class `BioeconomicParams`. MORE WORDS HERE.... Parameters from the Sea Bass fishery studied in @Tidbury2021 is included in the package and can be loaded from the `"seabass"`dataset.

# Running the model

## Specifying recruitment
* Different recruitment specifications.
* Making plots.


We can load in some parameter values
```{r load_data}
data("seabass")
```
Then we run it
```{r run_sim}
sim <- project(params, R = 1000, t_start= 2000, t_end = 2040)
```
Then we plot it
```{r fig.dim = c(7, 8)}
plot(sim)
```


# References
